---
title: "Chapter5"
author: "Melissa Wong"
date: \today
output: html_document
---

```{r results='hide', message=FALSE, warning=FALSE}
options("scipen" = 1, "digits" = 4)

library(tidyverse)
library(dagitty)
library(brms)
library(rethinking)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

# Practice Problems

5E1.  2, 3 and 4 are multiple linear regression models.

5E2.  $D_i = \alpha + \beta_L L + \beta_P P$

5E3.  $T_i = \alpha + \beta_F F + \beta_S S$ where $\beta_F, \beta_S > 0$

5E4. 1, 3, 4 and 5 are equivalent.

5M1.  

```{r results="hide", message=FALSE}
N <-50
df <- tibble(
  x_real = rnorm(N, 0, 1),
  x_spur = rnorm(N, x_real, 1),
  y = rnorm(N, 2 * x_real, 1)
)

mdl1 <- brm(y ~ x_spur, data=df)
mdl2 <- brm(y ~ x_real, data=df)
mdl3 <- brm(y ~ x_real + x_spur, data=df)
```

```{r}
fixef(mdl1)
fixef(mdl2)
fixef(mdl3)
```

5M2.

```{r results="hide", message=FALSE}
N <-50
df <- tibble(
  x_real = rnorm(N, 0, 1),
  x_spur = rnorm(N, x_real, 1),
  y = rnorm(N, x_real - x_spur, 1)
)

mdl1 <- brm(y ~ x_spur, data=df)
mdl2 <- brm(y ~ x_real, data=df)
mdl3 <- brm(y ~ x_real + x_spur, data=df)
```

```{r}
fixef(mdl1)
fixef(mdl2)
fixef(mdl3)
```

5M3. Higher divorce rates would also cause a higher rate of 2nd/3rd/etc. marriages.  So if we conditioned on rate of  re-marriages (or 1st marriages), then we could evaluate the effects of divorce rate and re-marriage rate on the overall marriage rate. 

5M4.

```{r}
mormon <- readRDS("Data_5M4.rds") %>%
  mutate(percent = as.numeric(sub("%","",Percent))) %>%
  select(-Percent)

data(WaffleDivorce)
divorce <- WaffleDivorce %>%
  select(Location, Loc, Marriage, 
         Divorce, MedianAgeMarriage) %>%
  left_join(mormon, by=c("Location" = "State")) %>%
  # standardize numeric variables
  mutate_if(is.numeric, scale)

```


