---
title: Chapter 8
author: "Melissa Wong"
date: \today
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r libraries, results='hide', message=FALSE, warning=FALSE}

library(tidyverse)
library(RColorBrewer)
library(rethinking)
```

```{r options}
options("scipen" = 1, "digits" = 4)

#knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(message = FALSE)
#knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(out.width = "50%")
knitr::opts_chunk$set(fig.align = "center")

# Set Default ggplot palette
options(ggplot2.discrete.color=brewer.pal(8, "Dark2"))
options(ggplot2.discrete.fill=brewer.pal(8, "Dark2"))
```

# 8E1

1. water temperature

2. work location

3. Pressing on the accelerator

# 8E2

Item 1 describes an interaction.

# 8E3

1. $Carmelize = \alpha + \beta_1 Temp + \beta_2 Dry + \beta_3 Temp * Dry$

2. $Speed = \alpha + \beta_1 Cyl + \beta_2 Injection$

3. $Belief = \alpha + \beta_1 parents + beta_2 friends$

4. $Intel = \alpha + \beta_1 social + \beta_2 appendage$

# 8M1

Under cold temp, blooms depend on both water and shade levels.

Under hot temp, there are no blooms regardless of water and shade levels.

So there are interactions between all 3 factors (temp, water and shade).

# 8M2

$\mu_T = \alpha_{[T]} + \beta_{W[T]} W + \beta_{S[T]} S + \beta_{WS[T]} WS$

where $\alpha_{[T]} = \beta_{W[T]} = \beta_{S[T]} = \beta_{WS[T]} = 0$ when $T=hot$

# 8M3

# 8M4

```{r}
data(tulips)

df <- tulips %>%
  mutate(blooms_std = blooms/max(blooms),
         water_cent = water - mean(water),
         shade_cent = shade - mean(shade),
         bed = as.factor(bed))
  
mdl8m4 <- quap(
  alist(
    blooms_std ~ dnorm(mu, sigma),
    mu <- a + bw * water_cent + bs*shade_cent + bws * water_cent * shade_cent,
    a ~ dnorm(0.5, 0.25),
    bw ~ dunif(0, 0.5), #dnorm(0, 0.25),
    bs ~ dunif(-0.5, 0), #dnorm(0, 0.25),
    bws ~ dnorm(0, 0.25),
    sigma ~ dexp(1)
  ),
  data=df
)

precis(mdl8m4)

# prior predictive simulation
prior <- extract.prior(mdl8m4, n=20)

grid <- data.frame(expand.grid(blooms_std = c(0, 0.5, 1), 
                               water_cent = c(-1, 0, 1), 
                               shade_cent = c(-1, 0, 1)))

blooms_pred <- link(mdl8m4, data=grid, post=prior, n=25)

cbind(grid, t(blooms_pred)) %>%
  pivot_longer(cols=-c("blooms_std", "water_cent", "shade_cent"),
               names_to="iter", values_to="blooms_prior") %>%
  ggplot() +
  geom_line(aes(x=water_cent, y=blooms_pred, group=iter),
            alpha=0.5) +
  facet_wrap(vars(shade_cent), ncol=3, labeller = label_both)

```

# 8H1

```{r}

# Interaction model w/bed as predictor
 m8H1 <- quap(
   alist(
     blooms_std ~ dnorm(mu, sigma),
     mu <- a[bed] + bw[bed]*water_cent + bs[bed]*shade_cent + 
       bws[bed] * water_cent * shade_cent,
     a[bed] ~ dnorm(0.5, 0.25),
     bw[bed] ~ dnorm(0, 0.25),
     bs[bed] ~ dnorm(0, 0.25),
     bws[bed] ~ dnorm(0, 0.25),
     sigma ~ dexp(1)
   ),
   data=df
 )
 
 precis(m8H1, depth=2)
 
```
# 8H2

```{r}
# Interaction model p256
 m8.5 <- quap(
   alist(
     blooms_std ~ dnorm(mu, sigma),
     mu <- a + bw*water_cent + bs*shade_cent + bws * water_cent * shade_cent,
     a ~ dnorm(0.5, 0.25),
     bw ~ dnorm(0, 0.25),
     bs ~ dnorm(0, 0.25),
     bws ~ dnorm(0, 0.25),
     sigma ~ dexp(1)
   ),
   data=df
 )

precis(m8.5)
```

```{r fig.height=10}
 compare(m8.5, m8H1)
 plot(coeftab(m8.5, m8H1))
```

The standard error for the difference is >> than the difference in WAIC between the two models; therefore adding _bed_ as a predictor doesn't significantly improve the model fit. Looking at the posterior distributions for the coefficients, there is significant overlap for _bs_, _bw_ and _bws_ for the three beds (i.e., we can't say the slopes differ significantly based on _bed_). The intercept for _bed_1_ is the only term whose 95% credible interval doesn't overlap at all with the intercepts for the other two beds; so the effect is _bed_1_ has slightly lower blooms, but not enough to significantly improve the model's predictions.

# 8M3

## Part a

```{r}
data(rugged)

df <- rugged %>%
  select(c(rgdppc_2000, rugged, cont_africa, country)) %>%
  mutate(log_gdp = log(rgdppc_2000)) %>%
  filter(complete.cases(.)) %>%
  mutate(log_gdp_std = log_gdp/mean(log_gdp),
         rugged_std = rugged/max(rugged),
         cid = ifelse(cont_africa == 1, 1, 2))

df %>%
  ggplot(mapping=aes(x=rugged_std, y=log_gdp_std)) +
  geom_point() +
  geom_text(aes(label=ifelse(log_gdp_std > 1.0 & rugged_std > 0.5,
                              as.character(country), ''))) +
  facet_wrap(vars(cid), labeller = label_both)
```

```{r}
m8.3 <- quap(
  alist(
    log_gdp_std ~ dnorm(mu, sigma),
    mu <- a[cid] + b[cid]*(rugged_std - 0.215),
    a[cid] ~ dnorm(1, 0.1),
    b[cid] ~ dnorm(0, 0.3),
    sigma ~ dexp(1)
  ),
  data=df
)

precis(m8.3, depth=2)
```


```{r}
df %>%
  mutate(waic = WAIC(m8.3, pointwise = TRUE)$penalty,
           psis = PSIS(m8.3, pointwise = TRUE)$k) %>%
  ggplot(mapping=aes(x=psis, y=waic)) +
  geom_point() +
  geom_text(aes(label=ifelse(waic > 0.5 & psis > 0.5,
                              as.character(country), ''))) +
  facet_wrap(vars(cid), labeller = label_both)
```
According to the plot above, none of the countries are overly influential (i.e., none have very large WAIC and/or PSIS values).

## Part b

```{r}
m8.3b <- quap(
  alist(
    log_gdp_std ~ dstudent(2, mu, sigma),
    mu <- a[cid] + b[cid]*(rugged_std - 0.215),
    a[cid] ~ dnorm(1, 0.1),
    b[cid] ~ dnorm(0, 0.3),
    sigma ~ dexp(1)
  ),
  data=df
)

precis(m8.3b, depth=2)
```

```{r}
plot(coeftab(m8.3, m8.3b))
```

There is not a substantial change in the results.

# 8H4

```{r}
data(nettle)

df <- nettle %>%
  mutate(log.lang.per.cap = log(num.lang / k.pop),
         log.area = log(area))

summary(df)

df %>%
  ggplot() +
  geom_point(mapping=aes(y=log.lang.per.cap, x=mean.growing.season))
```

## Part a

```{r}
mdl8H4a <- quap(
  alist(
    log.lang.per.cap ~ dnorm(mu, sigma),
    mu <- a + bm * mean.growing.season + ba * log.area,
    a ~ dnorm(-5, 3),
    bm ~ dnorm(0, 2),
    ba ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ), 
  data=df
)

# Prior predictive distribution
precis(mdl8H4a, prob=0.89)
```
The 89% credible interval for _bm_ is positive and doesn't include zero, so this is consistent with the hypothesis that _mean.growing.season_ is positively associated with the number of languages. 

## Part b

```{r}
mdl8H4b <- quap(
  alist(
    log.lang.per.cap ~ dnorm(mu, sigma),
    mu <- a +  bg * sd.growing.season + ba * log.area,
    a ~ dnorm(-5, 3),
    bg ~ dnorm(0, 2),
    ba ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ), 
  data=df
)

# Prior predictive distribution
precis(mdl8H4b)
```

The 89% credible interval for _bg_ includes zero, so we can't conclude anything about the association between _sd.growing.season_ and language diversity.

## Part c

```{r}
mdl8H4c <- quap(
  alist(
    log.lang.per.cap ~ dnorm(mu, sigma),
    mu <- a + bm * mean.growing.season + bg * sd.growing.season + 
      bgm * mean.growing.season * sd.growing.season  + ba * log.area,
    a ~ dnorm(-5, 3),
    bm ~ dnorm(0, 2),
    bg ~ dunif(0, 2),
    bgm ~ dnorm(0, 2),
    ba ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ), 
  data=df
)

# Prior predictive distribution
precis(mdl8H4c)
```

Including both terms and the interaction, now we conclude the following:

* _mean.growing.season_ is positively associated with language diversity

* the interaction between _mean.growing.season_ and _sd.growing.season_ is negatively associated with language diversity.
