---
title: "Chapter 6"
author: "Melissa Wong"
date: \today
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r results='hide', message=FALSE, warning=FALSE}
options("scipen" = 1, "digits" = 4)

library(tidyverse)
library(rethinking)
library(dagitty)
library(brms)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

# Practice Problems

## 6E1

1. Multicollinearity
2. Collider Bias
3. (Unobserved) Confounders

## 6E2

Simpson's Paradox is an example of confounder bias. Example is gender and college admissions.

```{r echo=FALSE}
d1 <- dagitty('dag{ 
  gender -> dept
  gender -> admitted <- dept
}')

plot(graphLayout(d1))
```

Berkson's paradox is an example of collider bias.

## 6E3

1. Fork: X <- Z -> Y (X is not a causal factor for Y but rather is correlated through Z)

2. Pipe: X -> Z -> (X is an indirect causal factor for Y)

3. Collider: X -> Z <- Y (X is not associated with Y unless conditioning on Z)

4. Descendant: X -> Z <- Y and Z -> D (X is not associated with Y unless conditioning on D)

## 6M1

```{r}
dag_6M1 <- dagitty("dag{ 
                  U [unobserved]
                  V [unobserved]
                  X -> Y
                  X <- U <- A -> C -> Y
                  U -> B <- C
                  C <- V -> Y
}")

plot(graphLayout(dag_6M1))
```

There are 5 paths from X to Y:

X - Y (open)

X - U - B - C - Y (closed because of B)

X - U - B - C - V - Y (closed because of B)

X - U - A - C - Y (open)

X - U - A _ C - V - Y (open)

We need to close the last two paths by conditioning on A.

```{r}
# Confirm
adjustmentSets(dag_6M1, exposure="X", outcome="Y")
```

# 6M2

```{r}
N <- 100
df <- data.frame(
  x <- rnorm(N),
  z <- 2*x + rnorm(N, 0, 0.5),
  y <- rnorm(N, z)
)
pairs(df)

mdl_6M2 <- quap(
  alist(
    y ~ dnorm(mu, sigma),
    mu <- a + b1 * x + b2 * z,
    a ~ dnorm(0, 1),
    b1 ~ dnorm(0, 1),
    b2 ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ),
  data = df
)

plot(precis(mdl_6M2))
```

This model does not exhibit multicollinearity.  Instead, what we can deduce from the plot above is that once we know $z$, knowing $x$ doesn't provide additional information about $y$.  This is consistent with our DAG $X \to Z \to Y$ that $x$ affects $y$ through $z$.

# 6M3

```{r include=FALSE}
dag_6M3 <- dagitty("dag{ 
                  X -> Y
                  X <- Z -> Y
                  Z <- A -> Y
}")

plot(graphLayout(dag_6M3))

adjustmentSets(dag_6M3, exposure="X", outcome="Y")
```


Upper left: condition on Z

Upper right: No need to condition for total causal effect

Bottom left: condition on Z and A

Bottom right: condition on A

# 6H1

```{r}
data("WaffleDivorce")

df <- WaffleDivorce %>%
  select(Loc, South, WaffleHouses, MedianAgeMarriage, Marriage, Divorce) %>%
  mutate(across(WaffleHouses:Divorce, scale))

dag_6H1 <- dagitty("dag{ 
                   WaffleHouses <- South -> MedianAge
                   South -> Marriage
                   MedianAge -> Marriage
                   MedianAge -> Divorce
                   Marriage -> Divorce
                   }")

plot(graphLayout(dag_6H1))

adjustmentSets(dag_6H1, exposure="WaffleHouses", outcome="Divorce")
```

```{r}
mdl1 <- quap(
  alist(
    Divorce ~ dnorm(mu, sigma),
    mu <- a + b1 *  WaffleHouses + b2 * South,
    a ~ dnorm(0, 1),
    b1 ~ dnorm(0, 1),
    b2 ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ),
  data=df
)

plot(precis(mdl1))
```

The 89% credible interval includes 0 for the effect of WaffleHouses.  However, the estimates for the intercept and both coefficients are all quite wide intervals suggesting this model isn't a great fit to the data.

# 6H2

Check if data is consistent with conditional independencies implied by the model

```{r}
impliedConditionalIndependencies(dag_6H1)
```


```{r}
mdl1 <- quap(
  alist(
    Divorce ~ dnorm(mu, sigma),
    mu <- a + b1 *  South + b2 * Marriage + b3 * MedianAgeMarriage,
    a ~ dnorm(0, 1),
    b1 ~ dnorm(0, 1),
    b2 ~ dnorm(0, 1),
    b3 ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ),
  data=df
)

plot(precis(mdl1))
```

Not consistent with the data, so try the following model instead.

```{r}
data("WaffleDivorce")

df <- WaffleDivorce %>%
  select(Loc, South, WaffleHouses, MedianAgeMarriage, Marriage, Divorce) %>%
  mutate(across(WaffleHouses:Divorce, scale))

dag_6H1b <- dagitty("dag{ 
                   WaffleHouses <- South -> MedianAge
                   South -> Marriage
                   South -> Divorce
                   MedianAge -> Marriage
                   MedianAge -> Divorce
                   }")

plot(graphLayout(dag_6H1b))

impliedConditionalIndependencies(dag_6H1b)
```

```{r}
mdl1 <- quap(
  alist(
    Divorce ~ dnorm(mu, sigma),
    mu <- a + b1 * Marriage + b2 * MedianAgeMarriage + b3 * South,
    a ~ dnorm(0, 1),
    b1 ~ dnorm(0, 1),
    b2 ~ dnorm(0, 1),
    b3 ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ),
  data=df
)

plot(precis(mdl1))
```
```{r}
mdl2 <- quap(
  alist(
    Divorce ~ dnorm(mu, sigma),
    mu <- a + b1 * South + b2 * WaffleHouses,
    a ~ dnorm(0, 1),
    b1 ~ dnorm(0, 1),
    b2 ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ),
  data=df
)

plot(precis(mdl2))
```
```{r}
mdl3 <- quap(
  alist(
    Marriage ~ dnorm(mu, sigma),
    mu <- a + b1 * South + b2 * WaffleHouses,
    a ~ dnorm(0, 1),
    b1 ~ dnorm(0, 1),
    b2 ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ),
  data=df
)

plot(precis(mdl3))
```
```{r}
mdl4 <- quap(
  alist(
    MedianAgeMarriage ~ dnorm(mu, sigma),
    mu <- a + b1 * South + b2 * WaffleHouses,
    a ~ dnorm(0, 1),
    b1 ~ dnorm(0, 1),
    b2 ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ),
  data=df
)

plot(precis(mdl4))
```

This 2nd DAG isn't quite right either unfortunately. There are likely other unobserved variables that should be in the model (maybe percentage of different religions?).

# 6H3

```{r}
data(foxes)

df <- foxes %>%
  mutate(across(avgfood:weight, scale))

mdl_6H3 <- quap(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + b * area,
    a ~ dnorm(0, 1),
    b ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ),
  data = df
)

precis(mdl_6H3)
```
The MAP estimate for total causal effect is very slightly positive, but the 89% credible interval includes 0 so we really can't say there is a causal effect of area on weight.

# 6H4

```{r}
mdl_6H4 <- quap(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + b * avgfood,
    a ~ dnorm(0, 1),
    b ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ),
  data = df
)

precis(mdl_6H4)
```

We don't need to condition on any variables to get the total causal effect of _avgfood_ on _weight_. Similarly, while the MAP estimate is slightly negative the 89% credible interal includes 0 so it does not appear there is any causal relationship between _avgfood_ and _weight_.

# 6H5

```{r}
mdl_6H5 <- quap(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + b1 * groupsize + b2 * avgfood,
    a ~ dnorm(0, 1),
    b1 ~ dnorm(0, 1),
    b2 ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ),
  data = df
)

precis(mdl_6H5)
```

We need to condition on _avgfood_ to correctly estimate the causal effect of _groupsize_.  Now the MAP estimate is -0.69 and the 89% credible interval is strictly negative, so increasing _groupsize_ would decrease _weight_.

Now the causal effect of _avgfood_ is positive.  Since there are two forward paths from _avgfood_ to _weight_, this suggests the the positive effect through _groupsize_ is essentially negated by the direct effect on _weight_.  In other words, the net effect of changing _avgfood_ is zero.


