---
title: "Chapter 2"
author: "Melissa Wong"
date: \today
output:
  pdf_document: default
---

```{r results='hide', message=FALSE, warning=FALSE}
options("scipen" = 1, "digits" = 4)

knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(out.width = "50%")
knitr::opts_chunk$set(fig.align = "center")

library(tidyverse)

library(rethinking)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```


# Globe Tossing Examples

## Grid Approximation
```{r}
N <- 20
p_grid <- seq(0, 1, length.out = N)
prior <- rep(1, N)
likelihood <- dbinom(6, size=9, prob=p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior/sum(unstd.posterior)

data.frame(x = p_grid, prior = prior, posterior=posterior) %>%
  pivot_longer(-x, names_to = "type", values_to = "prob") %>%
  ggplot(aes(x=x, y=prob, colour=type)) +
  geom_point() +
  geom_line() +
  labs(x="probability of water", y="posterior probability")

```

```{r}
N <- 20
p_grid <- seq(0, 1, length.out = N)
prior <- ifelse(p_grid < 0.5, 0, 1)
likelihood <- dbinom(6, size=9, prob=p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior/sum(unstd.posterior)

data.frame(x = p_grid, prior = prior, posterior=posterior) %>%
  pivot_longer(-x, names_to = "type", values_to = "prob") %>%
  ggplot(aes(x=x, y=prob, colour=type)) +
  geom_point() +
  geom_line() +
  labs(x="probability of water", y="posterior probability")
```
```{r}
N <- 20
p_grid <- seq(0, 1, length.out = N)
prior <- exp(-5*abs(p_grid - 0.5))
likelihood <- dbinom(6, size=9, prob=p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior/sum(unstd.posterior)

data.frame(x = p_grid, prior = prior, posterior=posterior) %>%
  pivot_longer(-x, names_to = "type", values_to = "prob") %>%
  ggplot(aes(x=x, y=prob, colour=type)) +
  geom_point() +
  geom_line() +
  labs(x="probability of water", y="posterior probability")
```

## Quadratic Approximation

```{r}
globe.qa <- quap(
  alist(
    W ~ dbinom(W+L, p),  # binomial likelihood
    p ~ dunif(0,1)       # uniform prior
  ),
  data=list(W=6, L=3))

# display quadratic approximation
precis(globe.qa)

```

```{r}
# analytic calculation
W <- 6
L <- 3
curve(dbeta(x, W+1, L+1), from=0, to=1)
curve(dnorm(x, 0.67, 0.16), lty=2, add=TRUE)
```

## MCMC

```{r}
N <- 1000
p <- rep(NA, N)
p[1] <- 0.5
W <- 6
L <- 3
for (i in 2:N)
{
  p_new <- rnorm(1, p[i-1], 0.1)
  if (p_new < 0) p_new <- abs(p_new)
  if (p_new > 1) p_new <- 2 - p_new
  q0 <- dbinom(W, W+L, p[i-1])
  q1 <- dbinom(W, W+L, p_new)
  p[i] <- ifelse(runif(1) < q1/q0, p_new, p[i-1])
}

dens(p, xlim=c(0,1))
curve(dbeta(x, W+1, L+1), lty=2, add=TRUE)
```

# Practice Problems

2E1 Pr(rain|Monday)

2E2 The probability it is Monday given it is raining

2E3 Pr(Monday|rain)

2E4 We don't know exactly what % of the globe is water, so we estimate it at 70%

2M1

```{r}
N <- 20
p_grid <- seq(0, 1, length.out = N)
prior <- rep(1, N)
likelihood <- dbinom(3, size=3, prob=p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior/sum(unstd.posterior)

data.frame(x = p_grid, prior = prior, posterior=posterior) %>%
  pivot_longer(-x, names_to = "type", values_to = "prob") %>%
  ggplot(aes(x=x, y=prob, colour=type)) +
  geom_point() +
  geom_line() +
  labs(x="probability of water", y="posterior probability")

```

```{r}
N <- 20
p_grid <- seq(0, 1, length.out = N)
prior <- rep(1, N)
likelihood <- dbinom(3, size=4, prob=p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior/sum(unstd.posterior)

data.frame(x = p_grid, prior = prior, posterior=posterior) %>%
  pivot_longer(-x, names_to = "type", values_to = "prob") %>%
  ggplot(aes(x=x, y=prob, colour=type)) +
  geom_point() +
  geom_line() +
  labs(x="probability of water", y="posterior probability")

```

```{r}
N <- 20
p_grid <- seq(0, 1, length.out = N)
prior <- rep(1, N)
likelihood <- dbinom(5, size=7, prob=p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior/sum(unstd.posterior)

data.frame(x = p_grid, prior = prior, posterior=posterior) %>%
  pivot_longer(-x, names_to = "type", values_to = "prob") %>%
  ggplot(aes(x=x, y=prob, colour=type)) +
  geom_point() +
  geom_line() +
  labs(x="probability of water", y="posterior probability")

```

2M3 Given $P[Land|Earth] = 0.3$ and $P[Land|Mars] = 1$, then 

\begin{align*}
  P[Earth|land] &= \frac{P[Earth, land]}{P[land]} \\
                &= \frac{0.3 * 0.5}{0.3*0.5 + 1*0.5} \\
                &= 0.2308
\end{align*}

2M4 Given 3 cards, 1 has two black side, 1 has 2 white sides and 1 has 1 black and 1 white.  If a card is selected and the black side is up, what is the probability the other side is black?

$$P[B/B | B] = \frac{P[B | B/B]P[B/B]}{P[B]} =\frac{1 * 1/3}{3/6} = \frac{2}{3}$$

2M5 Now suppose there are four cards: B/B,B/W, W/W, B/B.

$$P[B/B | B] = \frac{P[B | B/B]P[B/B]}{P[B]} =\frac{1 * 1/2}{5/8} = \frac{4}{5}$$