---
title: Chapter 8
author: "Melissa Wong"
date: \today
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r libraries, results='hide', message=FALSE, warning=FALSE}

library(tidyverse)
library(RColorBrewer)
library(rethinking)
```

```{r options}
options("scipen" = 1, "digits" = 4)

#knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(message = FALSE)
#knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(out.width = "50%")
knitr::opts_chunk$set(fig.align = "center")

# Set Default ggplot palette
options(ggplot2.discrete.color=brewer.pal(8, "Dark2"))
options(ggplot2.discrete.fill=brewer.pal(8, "Dark2"))
```

# 8E1

1. water temperature

2. work location

3. Pressing on the accelerator

# 8E2

Item 1 describes an interaction.

# 8E3

1. $Carmelize = \alpha + \beta_1 Temp + \beta_2 Dry + \beta_3 Temp * Dry$

2. $Speed = \alpha + \beta_1 Cyl + \beta_2 Injection$

3. $Belief = \alpha + \beta_1 parents + beta_2 friends$

4. $Intel = \alpha + \beta_1 social + \beta_2 appendage$

# 8M1

Under cold temp, blooms depend on both water and shade levels.

Under hot temp, there are no blooms regardless of water and shade levels.

So there are interactions between all 3 factors (temp, water and shade).

# 8M2

$\mu_T = \alpha_{[T]} + \beta_{W[T]} W + \beta_{S[T]} S + \beta_{WS[T]} WS$

where $\alpha_{[T]} = \beta_{W[T]} = \beta_{S[T]} = \beta_{WS[T]} = 0$ when $T=hot$

# 8M3

# 8M4

```{r}
data(tulips)

df <- tulips %>%
  mutate(blooms_std = blooms/max(blooms),
         water_cent = water - mean(water),
         shade_cent = shade - mean(shade))
  
mdl8m4 <- quap(
  alist(
    blooms_std ~ dnorm(mu, sigma),
    mu <- a + bw * water_cent + bs*shade_cent + bws * water_cent * shade_cent,
    a ~ dnorm(0.5, 0.25),
    bw ~ dunif(0, 0.5), #dnorm(0, 0.25),
    bs ~ dunif(-0.5, 0), #dnorm(0, 0.25),
    bws ~ dnorm(0, 0.25),
    sigma ~ dexp(1)
  ),
  data=df
)

precis(mdl8m4)

# prior predictive simulation
prior <- extract.prior(mdl8m4, n=20)

grid <- data.frame(expand.grid(blooms_std = c(0, 0.5, 1), 
                               water_cent = c(-1, 0, 1), 
                               shade_cent = c(-1, 0, 1)))

blooms_pred <- link(mdl8m4, data=grid, post=prior, n=25)

cbind(grid, t(blooms_pred)) %>%
  pivot_longer(cols=-c("blooms_std", "water_cent", "shade_cent"),
               names_to="iter", values_to="blooms_prior") %>%
  ggplot() +
  geom_line(aes(x=water_cent, y=blooms_pred, group=iter),
            alpha=0.5) +
  facet_wrap(vars(shade_cent), ncol=3, labeller = label_both)

```


