---
title: "Chapter5"
author: "Melissa Wong"
date: \today
output: html_document
---

```{r results='hide', message=FALSE, warning=FALSE}
options("scipen" = 1, "digits" = 4)

library(tidyverse)
library(dagitty)
library(brms)
library(rethinking)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

# Practice Problems

5E1.  2, 3 and 4 are multiple linear regression models.

5E2.  $D_i = \alpha + \beta_L L + \beta_P P$

5E3.  $T_i = \alpha + \beta_F F + \beta_S S$ where $\beta_F, \beta_S > 0$

5E4. 1, 3, 4 and 5 are equivalent.

5M1.  

```{r results="hide", message=FALSE}
N <-50
df <- tibble(
  x_real = rnorm(N, 0, 1),
  x_spur = rnorm(N, x_real, 1),
  y = rnorm(N, 2 * x_real, 1)
)

mdl1 <- brm(y ~ x_spur, data=df)
mdl2 <- brm(y ~ x_real, data=df)
mdl3 <- brm(y ~ x_real + x_spur, data=df)
```

```{r}
fixef(mdl1)
fixef(mdl2)
fixef(mdl3)
```

5M2.

```{r results="hide", message=FALSE}
N <-50
df <- tibble(
  x_real = rnorm(N, 0, 1),
  x_spur = rnorm(N, x_real, 1),
  y = rnorm(N, x_real - x_spur, 1)
)

mdl1 <- brm(y ~ x_spur, data=df)
mdl2 <- brm(y ~ x_real, data=df)
mdl3 <- brm(y ~ x_real + x_spur, data=df)
```

```{r}
fixef(mdl1)
fixef(mdl2)
fixef(mdl3)
```

5M3. Higher divorce rates would also cause a higher rate of 2nd/3rd/etc. marriages.  So if we conditioned on rate of  re-marriages (or 1st marriages), then we could evaluate the effects of divorce rate and re-marriage rate on the overall marriage rate. 

5M4.

```{r}
mormon <- readRDS("Data_5M4.rds") %>%
  mutate(percent = as.numeric(sub("%","",Percent))) %>%
  select(-Percent)

data(WaffleDivorce)
divorce <- WaffleDivorce %>%
  select(Location, Loc, Marriage, 
         Divorce, MedianAgeMarriage) %>%
  left_join(mormon, by=c("Location" = "State")) %>%
  # standardize numeric variables
  mutate_if(is.numeric, scale)

# Define model
mdl5.4 <- quap(
  alist(
    Divorce ~ dnorm(mu, sigma),
    mu <- a + b1 * Marriage + b2 * MedianAgeMarriage + b3 * percent,
    a ~ dnorm(0, 0.2),
    b1 ~ dnorm(0, 0.5),
    b2 ~ dnorm(0, 0.5),
    b3 ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = divorce
)

precis(mdl5.4)
```

```{r}
mu <- apply(link(mdl5.4), 2, mean)
ggplot(mapping=aes(x=mu, y=divorce$Divorce,
                         label=divorce$Loc)) +
  geom_text(color="blue") +
  geom_abline(mapping=aes(intercept=0, slope=1), 
                          linetype="dashed")
```
5M5.

```{r}
dag5M5 <- dagitty('dag{
  GasPrice -> Driving -> EatingOut -> Obesity
  GasPrice -> Driving -> Exercise -> Obesity
}')

plot(graphLayout(dag5M5))
```

If we have the predictors `Driving`, `EatingOut` and `Exercise` and we include all three in our model, then we can estimate the individual effects of the `EatingOut` and `Exercise` paths on Obesity.

5H1.

```{r}
dag5H1 <- dagitty('dag{
                  M -> A -> D
}')

print( impliedConditionalIndependencies(dag5H1))
```

```{r}
# Define model
mdl5H1 <- quap(
  alist(
    Divorce ~ dnorm(mu, sigma),
    mu <- a + b1 * Marriage + b2 * MedianAgeMarriage,
    a ~ dnorm(0, 0.2),
    b1 ~ dnorm(0, 0.5),
    b2 ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = divorce
)

precis(mdl5H1)
```

The 89% credible interval for `b1` includes 0 which is consistent with the model that Divorce and Marriage rates are independent conditional on MedianAgeMarriage.

5H2. 

```{r}
ppd <- sim(mdl5H1)

# Define sequence with 1/2 marriage rate
simdat <- data.frame(Marriage = divorce$Marriage/2)
# Simulate counterfactual data
scf <- sim(mdl5H1, data=simdat, 
         vars=c("MedianAgeMarriage", "Divorce"))

ggplot() +
  geom_text(mapping=aes(x=simdat$Marriage,
                        y=colMeans(scf$Divorce),
                        label=divorce$Loc,
                        color="Counterfactual")) +
  geom_text(mapping=aes(x=divorce$Marriage,
                        y=colMeans(ppd),
                        label=divorce$Loc,
                        color="Predicted"))
  labs(x="Marriage Rate",
       y="Divorce Rate")
```
5H3.

```{r}
data(milk)

df <- milk %>%
  filter(complete.cases(.)) %>%
  mutate(K = scale(kcal.per.g),
         N = scale(neocortex.perc),
         M = scale(log(mass)))

mdl5H3 <- quap(
  alist(
    ## M -> K <- N
    K ~ dnorm(mu, sigma),
    mu <- a + bN*N + bM*M,
    a ~ dnorm(0, 0.2),
    bN ~ dnorm(0, 0.5),
    bM ~ dnorm(0, 0.5),
    sigma ~ dexp(1),
    ## M -> N
    N ~ dnorm(muN, sigmaN),
    muN <- aN + bNM * M,
    aN ~ dnorm(0, 0.2),
    bNM ~ dnorm(0, 0.05),
    sigmaN ~ dexp(1)
  ),
  data = df
)

ppd <- sim(mdl5H3)

# Create manipulated values of M
simdat <- data.frame(M = df$M*2)
# Simulate counterfactual data
scf <- sim(mdl5H3, data=simdat, 
         vars=c("N", "K"))

ggplot() +
  geom_point(mapping=aes(x=simdat$M,
                        y=colMeans(scf$K),
                        color="Counterfactual")) +
  geom_point(mapping=aes(x=df$M,
                        y=colMeans(ppd),
                        color="Predicted")) +
  labs(x="log Body Mass", y="kcal per g")
```

