---
title: "Chapter 14"
author: "Melissa"
date: "10/10/2021"
output: html_document
---

```{r libraries, results='hide', message=FALSE, warning=FALSE}

library(tidyverse)
library(RColorBrewer)
#library(rethinking)
library(gridExtra)
library(tidybayes)
library(bayesplot)
#library(rstan)
library(dagitty)
library(here)
library(brms)
library(cmdstanr)
library(MASS)
```

```{r options}
options("scipen" = 1, "digits" = 4)

#knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(message = FALSE)
#knitr::opts_chunk$set(warning = FALSE)
#knitr::opts_chunk$set(out.width = "50%")
knitr::opts_chunk$set(fig.align = "center")

# Set Default ggplot palette
options(ggplot2.discrete.color=brewer.pal(8, "Dark2"))
options(ggplot2.discrete.fill=brewer.pal(8, "Dark2"))

set_cmdstan_path(here("../.cmdstanr/cmdstan-2.27.0"))
ncores <- min(4, parallel::detectCores()-1)

# Set brms backend to cmdstanr
options(brms.backend="cmdstanr")
```

# Section 14.1

```{r}
# Simulate population
a <- 3.5
b <- (-1)
sigma_a <- 1
sigma_b <- 0.5
rho <- (-0.7)
Mu <- c(a,b)
cov_ab <- sigma_a * sigma_b * rho
Sigma <- matrix(c(sigma_a^2, cov_ab, cov_ab, sigma_b^2), ncol=2)

# Alternate method for calculating covariance matrix
# sigmas <- c(sigma_a, sigma_b)
# Rho <- matrix(c(1, rho, rho, 1), ncol=2)
# Sigma <- diag(sigmas) %*% Rho %*% diag(sigmas)

N_cafes <- 20
set.seed(5)
vary_effects <- as.data.frame(mvrnorm(N_cafes, Mu, Sigma)) %>% 
  rename(a_cafe = V1, b_cafe = V2)

ggplot(data=vary_effects, mapping=aes(x=a_cafe, y=b_cafe)) +
  geom_point() +
  stat_ellipse(type="norm", level=0.25, alpha = 0.2) +
  stat_ellipse(type="norm", level=0.5, alpha = 0.2) +
  stat_ellipse(type="norm", level=0.75, alpha = 0.2) +
  stat_ellipse(type="norm", level=0.95, alpha = 0.2)


```

```{r}
# Simulation observations
N_visits <- 10
cafe_sim <- data.frame(cafe_id = rep(1:N_cafes, each=N_visits),
                 afternoon = rep(0:1, N_visits*N_cafes/2)) %>% 
  mutate(mu = vary_effects$a_cafe[cafe_id] + vary_effects$b_cafe[cafe_id] * afternoon) %>% 
  mutate(wait = rnorm(N_visits*N_cafes, mu, 0.5))

ggplot(cafe_sim) +
  geom_point(mapping=aes(x=cafe_id, y=wait, color=factor(afternoon)))
```
```{r results="hide"}
# Varying slopes model
cafe_prior <- c(set_prior("normal(5,2)", class="Intercept"),
                set_prior("normal(-1, 0.5)", class="b"),
                set_prior("exponential(1)", class="sd"),
                set_prior("exponential(1)", class="sigma"),
                set_prior("lkj(2)", class="cor"))

cafe_mdl <- brm(wait ~ 1 + afternoon + (1 + afternoon | cafe_id),
                data = cafe_sim,
                family=gaussian,
                chains=4,
                cores=ncores,
                file_refit = "on_change",
                file = here("mdl14-1.rds"))
```

```{r}
summary(cafe_mdl)
```

```{r eval=FALSE}
#get_variables(cafe_mdl)
#stancode(cafe_mdl)
#str(make_standata(wait ~ 1 + afternoon + (1 + afternoon | cafe_id),
#                  data = cafe_sim))
```

NOTES: 

* variables starting with `b_` are population level effects

** `b_Intercept` is population intercept

** `b_afternoon` is the population slope for _afternoon_ variable

* variables starting with `sd_` are group-level standard deviations

* `cor[Intercept, afternoon]` is the group-level correlation coefficient

* variables starting with `z_` are _standardized_ group level effects

* variables starting with `r_` are _actual_ group level effects

** `r_cafe_id[#, Intercept]` is delta from `b_Intercept` for cafe #

** `r_cafe_id[#, afternoon]` is delta from `b_afternoon` for cafe #

** `r_1[cafe#, 1]` is the same as `r_cafe_id[#, Intercept]` and `r_1[cafe#, 1]` as `r_cafe_id[#, afternoon]`

* `sigma` is the residual standard deviation

```{r}
mcmc_rank_overlay(cafe_mdl,
                  regex_pars = c("^b_", "^sd_", "^cor_", "sigma"))


mcmc_plot(cafe_mdl, type="intervals",
          pars=c("^b_", "^sd_", "^cor_", "sigma"),
          prob=0.5, prob_outer=0.95,
          point_est="median")
```

```{r}
# Random effects coefficients

# Manually construct coefficients by adding population and group parameters
post <- posterior_samples(cafe_mdl) %>% 
  dplyr::select(starts_with("b_") | starts_with("r_cafe_id")) 

# Add population-level intercept/slope 
cafe_intercepts <- (post[,1] + post[,3:22]) %>% 
  apply(2, function(x) quantile(x, probs=c(0.025, 0.5, 0.975))) %>% 
  t() %>% 
  as.data.frame()

cafe_slopes <- (post[,2] + post[,23:42]) %>% 
  apply(2, function(x) quantile(x, probs=c(0.025, 0.5, 0.975))) %>% 
  t() %>% 
  as.data.frame()

# The easy way to get random effects (i.e., group) coefficients
coefs <- coef(cafe_mdl)

# The easy way to get fixed effects (i.e., population) coefficients
fixef(cafe_mdl)
```

```{r}


ggplot(data=vary_effects, mapping=aes(x=a_cafe, y=b_cafe)) +
  geom_point() +
  geom_point(mapping=aes(x=cafe_intercepts$`50%`, y=cafe_slopes$`50%`),
             color="blue") +
  stat_ellipse(type="norm", level=0.25, alpha = 0.2) +
  stat_ellipse(type="norm", level=0.5, alpha = 0.2) +
  stat_ellipse(type="norm", level=0.75, alpha = 0.2) +
  stat_ellipse(type="norm", level=0.95, alpha = 0.2)
```

