---
title: "HW1"
author: "Melissa Wong"
date: "4/14/2021"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Problem 1

```{r problem1}
N <- 25
df1 <- tibble(p = seq(0, 1, length.out=N),
       prior = rep(1,N)) %>%
  mutate(likelihood = dbinom(4, 15, p),
          post = likelihood * prior) 
df1 %>%
  ggplot() +
  geom_line(mapping=aes(x=p, y=prior, color="prior")) +
  geom_line(mapping=aes(x=p, y=post/sum(post), color="posterior"))
```

# Problem 2

```{r problem2}
N <- 25
df2 <- tibble(p = seq(0, 1, length.out=N)) %>%
  mutate(prior = ifelse(p<0.5, 0, 1),
         likelihood = dbinom(4, 15, p),
          post = likelihood * prior) 
df2 %>%
  ggplot() +
  geom_line(mapping=aes(x=p, y=prior, color="prior")) +
  geom_line(mapping=aes(x=p, y=post/sum(post), color="posterior"))
```

This choice of prior effectively means values of $p<0.5$ are not possible.  Since our observed data is $\hat{p} = 4/15 < 0.5$, then it follows that the posterior distribution of $p$ has most of the probability mass concentrated around 0.5 which is the smallest possible value consistent with the prior.

# Problem 3

```{r}
samples <- sample(df2$p, prob=df2$post/sum(df2$post), 
                  size=1e4, replace=TRUE)
rethinking::PI(samples)
rethinking::HPDI(samples)
```

The percentile interval is wider because it assumes the distribution is symmetric--which in this case is not true.
