---
title: "HW3"
author: "Melissa Wong"
date: \today
output: pdf_document
---
```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rethinking)
library(dagitty)
library(brms)
library(RColorBrewer)

# mc-stan options
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

# Setup

```{r}
data(foxes)

df <- foxes %>%
  mutate(across(avgfood:weight, scale))

dag_fox <- dagitty('dag{
  area -> avgfood
  avgfood -> groupsize
  avgfood -> weight <- groupsize
}')

plot(graphLayout(dag_fox))

impliedConditionalIndependencies(dag_fox)
```

# Problem 1

To get the total causal effect of area on weight, simply regress weight on area.

```{r}
# Confirm no adjustment needed
adjustmentSets(dag_fox, exposure="area", outcome="weight")
```

```{r}
# Choose priors
# Fit model
mdl_fox <- quap(
  alist(
    # area -> weight
    weight ~ dnorm(mu_weight, sigma_weight),
    mu_weight <- a + b * area,
    a ~ dnorm(0, 1),
    b ~ dnorm(0, 1),
    sigma_weight ~ dexp(1)
  ), data=df
)

precis(mdl_fox)
```
# Problem 2

```{r}
mdl2 <- quap(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + b * avgfood,
    a ~ dnorm(0, 1),
    b ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ),
  data = df
)

precis(mdl2)
```

We don't need to condition on any variables to get the total causal effect of _avgfood_ on _weight_. Similarly, while the MAP estimate is slightly negative the 89% credible interal includes 0 so it does not appear there is any causal relationship between _avgfood_ and _weight_.

# Problem 3

```{r}
mdl3 <- quap(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + b1 * groupsize + b2 * avgfood,
    a ~ dnorm(0, 1),
    b1 ~ dnorm(0, 1),
    b2 ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ),
  data = df
)

precis(mdl3)
```

We need to condition on _avgfood_ to correctly estimate the causal effect of _groupsize_.  Now the MAP estimate is -0.69 and the 89% credible interval is strictly negative, so increasing _groupsize_ would decrease _weight_.

Now the causal effect of _avgfood_ is positive.  Since there are two forward paths from _avgfood_ to _weight_, this suggests the the positive effect through _groupsize_ is essentially negated by the direct effect on _weight_.  In other words, the net effect of changing _avgfood_ is zero.


