---
title: Assignment 6
author: "Melissa Wong"
date: \today
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r libraries, results='hide', message=FALSE, warning=FALSE}

library(tidyverse)
library(RColorBrewer)
library(rethinking)
library(gridExtra)
library(tidybayes)
library(bayesplot)
library(rstan)
library(cmdstanr)
library(dagitty)
library(here)
library(brms)

```

```{r options}
options("scipen" = 1, "digits" = 4)

#knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(message = FALSE)
#knitr::opts_chunk$set(warning = FALSE)
#knitr::opts_chunk$set(out.width = "50%")
knitr::opts_chunk$set(fig.align = "center")

# Set Default ggplot palette
options(ggplot2.discrete.color=brewer.pal(8, "Dark2"))
options(ggplot2.discrete.fill=brewer.pal(8, "Dark2"))
```

# Problem 1
 
```{r}
data("NWOGrants")

G <- dagitty("dag{ 
             awards <- gender -> discipline;
             discipline -> awards;
             }")

plot(graphLayout(G))

# adjustmentSets(G, outcome="awards", exposure="gender",
#                effect="direct")
```
$$ y_i \sim Binom(N, p_i)$$
$$logit(p_i) = \alpha_{gender[i]}$$

$$\alpha_{gender[i]} \sim N(0,1)$$
```{r results='hide'}
mdl1 <- brm(awards | trials(applications) ~ gender,
           data=NWOGrants,
           family=binomial(link="logit"),
           chains=4,
           file_refit = "on_change",
           file = here("homework/mdl6-1a.rds"),
           backend="cmdstanr")

#stancode(mdl)
```


```{r}
plot(mdl1, combo=c("dens", "rank_overlay"))

summary(mdl1)
```

```{r results='hide'}
mdl2 <- brm(awards | trials(applications) ~ gender + discipline,
           data=NWOGrants,
           family=binomial(link="logit"),
           chains=4,
           file_refit = "on_change",
           file = here("homework/mdl6-1b.rds"),
           backend = "cmdstanr")

#stancode(mdl2)
```

```{r}
mcmc_rank_overlay(mdl2)

summary(mdl2)
```

The total causal effect of _gender_ on receiving an award is that the odds are ~1.2 times ([1.01, 1.5] 95% credible interval) higher for males vs females.

The 95% credible interval for the _direct_ causal effect of _gender_ on receiving an award is [0.9324, 1.42] which includes 1, so in fact gender does not directly affect the odds of receiving an award. However, some departments are much more likely than others to receive awards.  If we wanted to increase the number of awards to females, the we would want to recruit more more females into the Chemical Sciences, Physics, Physical Sciences and Earth & Life Sciences departments.


```{r include=FALSE, eval=FALSE}
# rstan equivalent to above

# Total causal effect of gender
mdl1 <- stan(file=here("homework/m6-1.stan"), 
             data=list(N = nrow(NWOGrants),
                 applications = NWOGrants$applications,
                 awards = NWOGrants$awards,
                 gender = as.numeric(NWOGrants$gender),
                 Ngender = length(unique(NWOGrants$gender))),
             chains=4)

# Example of how to do predictions without refitting stan model
# Too much work--either use brms or do predictions in R 
# using stanfit posteriors
params <- extract(mdl1, pars="alpha")
md1_pred <- stan(file=here("homework/m6-1_pred.stan"), 
             data=list(N = nrow(NWOGrants),
                 applications = NWOGrants$applications,
                 gender = as.numeric(NWOGrants$gender),
                 Ngender = length(unique(NWOGrants$gender)),
                 Nsamples = nrow(params$alpha),
                 alpha = params$alpha),
             algorithm = "Fixed_param",
             chains=1)

print(mdl1)

mcmc_rank_overlay(mdl1)

mdl1_draws <- mdl1 %>%
  gather_draws(alpha[i]) %>%
  mutate(gender = ifelse(i==1, "male", "female")) %>%
  ungroup() %>%
  select(-i)

contrast <- mdl1_draws %>%
  pivot_wider(names_from=gender, values_from=.value) %>%
  mutate(delta = male - female)

contrast %>%
  ggplot() +
  geom_density(aes(x=delta))
```

# Problem 2

```{r}
G2 <- dagitty("dag{ 
             awards <- gender -> discipline;
             discipline -> awards;
             discipline <- U -> awards;
             U [unobserved];
             }")

plot(graphLayout(G2))

impliedConditionalIndependencies(G2)

adjustmentSets(G2, outcome="awards", exposure="gender",
               effect="total")
```

Conditioning on discipline "opens" the backdoor path through the unobserved variable, so it no longer provides an un-confounded estimate of the direct effect of gender.

# Problem 3

```{r}
library(MASS)
data(eagles)

E <- dagitty("dag{ 
             Success <- PirateAge -> PirateSize;
             PirateSize -> Success;
             VictimSize -> Success;
             }")

plot(graphLayout(E))
```

Fit three models:

1. Total causal effect of pirate age

2. Assume no interaction between pirate and victim size

3. Allow for interaction between pirate and victim size

Note: `mdl3b` and `mdl3c` are sensitive to the choice of prior, most likely due to the small amount of data. In fact `mdl3c` has poor sample size in the tails when using a flat prior--a weakly informative prior dramatically improves sampling in the tails. 

## Pirate Age, Victim Size only model

```{r results="hide"}
mdl3a <- brm(y | trials(n) ~ A + V,
              data=eagles,
              family=binomial(link="logit"),
             prior=set_prior(
               "normal(0,2)",
               class="b"
             ),
              chains=4,
              file_refit = "on_change",
            file = here("homework/mdl3a.rds"),
            backend="cmdstanr")
```


```{r}
plot(mdl3a, combo=c("dens", "rank_overlay"))

summary(mdl3a)
```

## All features, no interaction Model

```{r results="hide"}
mdl3b <- brm(y | trials(n) ~ A + V + P,
              data=eagles,
              family=binomial(link="logit"),
             prior=set_prior(
               "normal(0,2)",
               class="b"
             ),
              chains=4,
              file_refit = "on_change",
            file = here("homework/mdl3b.rds"),
            backend="cmdstanr")
```


```{r}
plot(mdl3b, combo=c("dens", "rank_overlay"))

summary(mdl3b)
```

## All features, size interaction Model

```{r results="hide"}
mdl3c <- brm(y | trials(n) ~ A + V * P,
             data=eagles,
             family=binomial(link="logit"),
             prior=set_prior(
               "normal(0,2)",
               class="b"
             ),
             chains=4,
             file_refit = "on_change",
             file = here("homework/mdl3c.rds"),
             backend="cmdstanr")
```


```{r}
plot(mdl3c, combo=c("dens", "rank_overlay"))

summary(mdl3c)
```

```{r include=FALSE, eval=FALSE}

get_variables(mdl3c)

# Get posterior predictions
add_epred_draws(eagles, mdl3c) %>%
  median_qi()

add_epred_draws(eagles, mdl3c) %>%
  median_hdi()

# expected posterior (n * p)
epred <- add_epred_draws(eagles, mdl3c)
# log-odds scale
linpred <- add_linpred_draws(eagles, mdl3c)
# posterior predicted (integers)
preds <- add_predicted_draws(eagles, mdl3c)
```


## Summary

```{r}
mdl3c <- add_criterion(mdl3c, "waic")
mdl3b <- add_criterion(mdl3b, "waic")
mdl3a <- add_criterion(mdl3a, "waic")
loo_compare(mdl3c, mdl3b, mdl3a, criterion = "waic")
```

Comparing the models with WAIC


```{r}
mdl3c <- add_criterion(mdl3c, "loo")
mdl3b <- add_criterion(mdl3b, "loo")
mdl3a <- add_criterion(mdl3a, "loo")
loo_compare(mdl3c, mdl3b, mdl3a, criterion = "loo")
```
We can see that `mdl3b` and `mdl3c` fit the data better (unsurprising since they include additional covariates).  However, there isn't a significant different between `mdl3b` and `mdl3c`--also not surprising since the 95% credible for the interaction term in `mdl3c` is squarely centered over 0.  Based on that, I would say the interaction between pirate and victim size is negligible.

The odds of success are about 1/3 for an immature pirate compared to a mature pirate, 54x greater when the victim is small vs large and about 1/36 when the pirate is small.
