---
title: Rounding & Truncating
date: \today
output:
  html_document:
    df_print: paged
  pdf_document: default
---

# Description

This file looks at the effects of rounding and truncation on parameter estimates for a simple linear model

$$y = \alpha + x\beta$$

There are three simulated data sets:

1.  y_true = raw data

2.  y_round = round(y_true) for a random subset of the data

3.  y_trunc = trunc(y_true) for a random subset of the data

Similarly there are three different models: *reference*, *rounded*, *truncated*.

A total of seven fits are performed:

1.  y_true with *reference* (This is the "reference" fit against which the others are compared).

2-4. y_round with *reference*, *rounded*, and *truncated*

5-7. y_trunc with *reference*, *rounded*, and *truncated*

# Setup

```{r libraries, results='hide', message=FALSE, warning=FALSE}

library(tidyverse)
library(RColorBrewer)
library(brms)
library(here)
library(bayesplot)
library(tidybayes)
#library(cmdstanr)
```

```{r options}
options("scipen" = 1, "digits" = 4)

#knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(message = FALSE)
#knitr::opts_chunk$set(warning = FALSE)
#knitr::opts_chunk$set(out.width = "50%")
#knitr::opts_chunk$set(fig.align = "center")

# Set Default ggplot palette
options(ggplot2.discrete.color=brewer.pal(8, "Dark2"))
options(ggplot2.discrete.fill=brewer.pal(8, "Dark2"))

ncores <- min(4, parallel::detectCores()-1)

# Set brms backend to cmdstanr
# options(brms.backend="cmdstanr")

set.seed(1234)
```

# Simulate Data

```{r message=FALSE}
N <- 1000
alpha_true = -1
beta_true = 1
sigma_true = 1

df <- data.frame(x = rnorm(N, 0, 5)) %>%
  mutate(y_true = alpha_true + beta_true*x + rnorm(N, 0, sigma_true),
         y_round = y_true,
         y_trunc = y_true,
         iflag = 0
  )

idx <- which(-5 < df$x & df$x < 20, arr.ind = TRUE)
sidx <- sample(idx, length(idx) %/% 2, replace=FALSE)

df$y_round[sidx] <- round(df$y_round[sidx])
df$y_trunc[sidx] <- trunc(df$y_trunc[sidx])
df$iflag[sidx] <- 1

df %>%
  select(-iflag) %>% 
  pivot_longer(-starts_with("x"), names_to="data") %>%
  ggplot() +
  geom_point(aes(x=x, y=value), alpha=0.2) +
  facet_wrap(~data)

df %>%
  select(-iflag) %>% 
  pivot_longer(-starts_with("x"), names_to="data") %>%
ggplot() +
  geom_histogram(aes(x=value, fill=data), alpha=0.8) +
  facet_wrap(~data)
```

# Reference Model

Fit `y_true` linear model as reference.

```{r results="hide"}

# Fit model
mdl1 <- brm(y_true ~ 1 + x,
            prior = c(set_prior("normal(0, 5)", class="Intercept"),
                      set_prior("exponential(1)", class="sigma")),
            data=df,
            family=gaussian(),
            chains=4,
            cores=ncores,
            file_refit = "on_change",
            file = here("rounding_example/mixed_mdl1.rds"))
```

```{r}
summary(mdl1)
```

```{r}
mcmc_rank_overlay(mdl1)
```

```{r}

# Save posterior draws for comparison later
mdl1_draws <- mdl1 %>%
  #gather_draws(alpha, beta[i], sigma) %>%
  gather_draws(b_Intercept, b_x, sigma) %>%
  mutate(model_id = "reference")
```


# Define custom brms functions

## Define Rounded Likelihood

```{r}
rounded_normal <- custom_family(
  "rounded_normal", 
  dpars = c("mu", "sigma"),
  lb = c(NA, 0),
  links = c("identity"), 
  type = "real",
  vars = "vint1[n]"
)

stan_funs_rounded_normal <- "
  real rounded_normal_lpdf(real y, real mu, real sigma, int iflag) {
    // Likelihood
    if (iflag == 1) {
      return log(Phi((y + 0.5 - mu)/sigma) - Phi((y - 0.5 - mu)/sigma));
    }
    else {
      return normal_lpdf(y | mu, sigma);
    }
  }
  real rounded_normal_rng(real mu, real sigma, int iflag) {
    if (iflag == 1) {
      return round(normal_rng(mu, sigma));    }
    else {
      return normal_rng(mu, sigma);
    }
    
  }
"

posterior_predict_rounded_normal <- function(i, prep, ...) {
  mu <- prep$dpars$mu[, i]
  sigma <- prep$dpars$sigma
  iflag <- prep$data$vint1[i]
  rounded_normal_rng(mu, sigma, iflag)
}

posterior_epred_rounded_normal <- function(prep) {
  mu <- prep$dpars$mu
}

log_lik_rounded_normal <- function(i, prep) {
  mu <- prep$dpars$mu[, i]
  sigma <- prep$dpars$sigma
  iflag <- prep$data$vint1[i]
  y <- prep$data$Y[i]
  rounded_normal_lpdf(y, mu, sigma, iflag)
}
```

## Define Truncated Likelihood

```{r}
truncated_normal <- custom_family(
  "truncated_normal", 
  dpars = c("mu", "sigma"),
  lb = c(NA, 0),
  links = c("identity"), 
  type = "real",
  vars = "vint1[n]"
)

stan_funs_truncated_normal <- "
  real truncated_normal_lpdf(real y, real mu, real sigma, int iflag) {
  
    // Likelihood
    if (iflag == 1) {
      if(y < 0.0) {
       return log(Phi((y - mu) / sigma)
          - Phi((y - 1.0 - mu) / sigma));     
      }
      else {
        return log(Phi((y + 1.0 - mu) / sigma)
            - Phi((y - mu) / sigma));
      }    
    }
    else {
      return normal_lpdf(y | mu, sigma);
    }
    // Likelihood
    
  }
  
  real truncated_normal_rng(real mu, real sigma, int iflag) {
    if (iflag == 1) {
        return trunc(normal_rng(mu, sigma));

    }
    else {
        return normal_rng(mu, sigma);
    }
  }
"

posterior_predict_truncated_normal <- function(i, prep, ...) {
  mu <- prep$dpars$mu[, i]
  sigma <- prep$dpars$sigma
  iflag <- prep$data$vint1[i]
  truncated_normal_rng(mu, sigma, iflag)
}

posterior_epred_truncated_normal <- function(prep) {
  mu <- prep$dpars$mu
}

log_lik_truncated_normal <- function(i, prep) {
  mu <- prep$dpars$mu[, i]
  sigma <- prep$dpars$sigma
  iflag <- prep$data$vint1[i]
  y <- prep$data$Y[i]
  truncated_normal_lpdf(y, mu, sigma, iflag)
}
```

# Rounded Data

## Incorrect Model (Reference)

First, fit `y_round` data to the reference model to see effect of *not* accounting for the rounding.

```{r results="hide"}
# Fit model
mdl2 <- brm(y_round ~ 1 + x,
            prior = c(set_prior("normal(0, 5)", class="Intercept"),
                      set_prior("exponential(1)", class="sigma")),
            data=df,
            family=gaussian(),
            chains=4,
            cores=ncores,
            file_refit = "on_change",
            file = here("rounding_example/mixed_mdl2.rds"))
```

```{r}
print(mdl2)
```

```{r}
mcmc_rank_overlay(mdl2)
```

```{r}
# Save posterior draws for comparison later
mdl2_draws <- mdl2 %>% 
  gather_draws(b_Intercept, b_x, sigma) %>%
  mutate(model_id = "round_reference")
```

## Correct Model

Next, fit `y_round` data to the correct model which accounts for rounding.

```{r results="hide"}
# Fit model
mdl3 <- brm(y_round | vint(iflag) ~ 1 + x,
            prior = c(set_prior("normal(0, 5)", class="Intercept"),
                      set_prior("exponential(1)", class="sigma")),
            data=df,
            family=rounded_normal,
            stanvars = stanvar(scode=stan_funs_rounded_normal,
                               block="functions"),
            #stan_funs = stan_funs_rounded_normal,
            chains=4,
            cores=ncores,
            file_refit = "on_change",
            file = here("rounding_example/mixed_mdl3.rds"))

# Expose stan functions so we can use post-processing functions
# like loo and pp_check
expose_functions(mdl3, vectorize = TRUE)
```

```{r}
print(mdl3)
```

```{r}
mcmc_rank_overlay(mdl3)
```

```{r}
# Save posterior draws for comparison later
mdl3_draws <- mdl3 %>% 
  gather_draws(b_Intercept, b_x, sigma) %>%
  mutate(model_id = "round_round")
```

## Incorrect Model (Truncated)

First, fit `y_round` data to the truncated model to see effect of incorrect model.

```{r results="hide"}
# Fit model
mdl7 <- brm(y_round | vint(iflag) ~ 1 + x,
            prior = c(set_prior("normal(0, 5)", class="Intercept"),
                      set_prior("exponential(1)", class="sigma")),
            data=df,
            family=truncated_normal,
            stanvars = stanvar(scode=stan_funs_truncated_normal,
                               block="functions"),
            #stan_funs = stan_funs_rounded_normal,
            chains=4,
            cores=ncores,
            file_refit = "on_change",
            file = here("rounding_example/mixed_mdl7.rds"))

# Expose stan functions so we can use post-processing functions
# like loo and pp_check
expose_functions(mdl7, vectorize = TRUE)
```

```{r}
print(mdl7)
```

```{r}
mcmc_rank_overlay(mdl7)
```

```{r}
# Save posterior draws for comparison later
mdl7_draws <- mdl7 %>% 
  gather_draws(b_Intercept, b_x, sigma) %>%
  mutate(model_id = "round_trunc")
```


## Compare

```{r}
loo(mdl2, mdl3, mdl7)
```

# Truncated Data

## Incorrect Model (Reference)

Fit `y_trunc` data to the reference model to see effect of *not* accounting for the truncating.

```{r results="hide"}
# Fit model
mdl4 <- brm(y_trunc  ~ 1 + x,
            prior = c(set_prior("normal(0, 5)", class="Intercept"),
                      set_prior("exponential(1)", class="sigma")),
            data=df,
            family=gaussian(),
            chains=4,
            cores=ncores,
            file_refit = "on_change",
            file = here("rounding_example/mixed_mdl4.rds"))
```

```{r}
print(mdl4)
```

```{r}
mcmc_rank_overlay(mdl4)
```

```{r}
# Save posterior draws for comparison later
mdl4_draws <- mdl4 %>% 
  gather_draws(b_Intercept, b_x, sigma) %>%
  mutate(model_id = "trunc_reference")
```

## Correct Model



Next, fit `y_trunc` data to the correct model which accounts for truncation.

```{r results="hide"}
# Fit model
mdl5 <- brm(y_trunc | vint(iflag) ~ 1 + x,
            prior = c(set_prior("normal(0, 5)", class="Intercept"),
                      set_prior("exponential(1)", class="sigma")),
            data=df,
            family=truncated_normal,
            stanvars = stanvar(scode=stan_funs_truncated_normal,
                               block="functions"),
            #stan_funs = stan_funs_rounded_normal,
            chains=4,
            cores=ncores,
            file_refit = "on_change",
            file = here("rounding_example/mixed_mdl5.rds"))
```

```{r}
print(mdl5)
```

```{r}
mcmc_rank_overlay(mdl5)
```

```{r}
# Save posterior draws for comparison later
mdl5_draws <- mdl5 %>% 
  gather_draws(b_Intercept, b_x, sigma) %>%
  mutate(model_id = "trunc_trunc")
```

## Incorrect Model (Rounding)

Next, fit `y_trunc` data to the model which assumes rounding.

```{r results="hide"}
# Fit model
mdl6 <- brm(y_trunc | vint(iflag) ~ 1 + x,
            prior = c(set_prior("normal(0, 5)", class="Intercept"),
                      set_prior("exponential(1)", class="sigma")),
            data=df,
            family=rounded_normal,
            stanvars = stanvar(scode=stan_funs_rounded_normal,
                               block="functions"),
            #stan_funs = stan_funs_rounded_normal,
            chains=4,
            cores=ncores,
            file_refit = "on_change",
            file = here("rounding_example/mixed_mdl6.rds"))
```

```{r}
print(mdl6)
```

```{r}
mcmc_rank_overlay(mdl6)
```

```{r}
# Save posterior draws for comparison later
mdl6_draws <- mdl6 %>% 
  gather_draws(b_Intercept, b_x, sigma) %>%
  mutate(model_id = "trunc_round")
```

## Compare

```{r}
loo(mdl4, mdl5, mdl6)
```

# Compare Models

```{r}
# Compare model fits
rbind(mdl1_draws, mdl2_draws,
      mdl3_draws, mdl4_draws,
      mdl5_draws, mdl6_draws,
      mdl7_draws) %>%
  group_by(.variable, model_id) %>%
  summarize(mean = mean(.value),
            lower = quantile(.value, probs=0.025),
            upper = quantile(.value, probs = 0.975),
            .groups="drop") %>%
  mutate(model_id = fct_relevel(model_id, "reference")) %>%
  # arrange(.variable, model_id) %>%
  # knitr::kable()
  ggplot(aes(x=mean, xmin=lower, xmax=upper,
             y=model_id, color=model_id),) +
  geom_errorbarh(aes(height=0.1)) +
  geom_point() +
  facet_wrap(~.variable, scales="free_x") +
  labs(x="Posterior 95% Interval") +
  geom_vline(data = data.frame(xint=c(alpha_true, beta_true, sigma_true), 
                               .variable=c("b_Intercept", "b_x", "sigma")),
             aes(xintercept = xint), color="red",
             linetype="dashed") +
  theme(axis.text.x = element_text(angle = 45)) +
  theme(legend.position = "none")
```

The red dashed lines are the true parameter values. When there is a mixture of truncated/rounded data, it's less critical to use the correct model when the proportion of truncated/rounded data is low (in this example approximately `r mean(df$iflag)`).   
