---
title: Rounding & Truncating
date: \today
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r libraries, results='hide', message=FALSE, warning=FALSE}

library(tidyverse)
library(RColorBrewer)
library(rstan)
library(here)
library(bayesplot)
library(tidybayes)
```

```{r options}
options("scipen" = 1, "digits" = 4)

#knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(message = FALSE)
#knitr::opts_chunk$set(warning = FALSE)
#knitr::opts_chunk$set(out.width = "50%")
#knitr::opts_chunk$set(fig.align = "center")

# Set Default ggplot palette
options(ggplot2.discrete.color=brewer.pal(8, "Dark2"))
options(ggplot2.discrete.fill=brewer.pal(8, "Dark2"))
```

# Simulate Data

```{r message=FALSE}
N <- 1000
alpha_true = -1
beta_true = 1
sigma_true = 1

df <- data.frame(x = rnorm(N, 0, 5)) %>%
  mutate(y_true = alpha_true + beta_true*x + rnorm(N, 0, sigma_true),
    y_round = round(y_true),
    y_trunc = trunc(y_true)
  )

df %>%
  pivot_longer(-starts_with("x"), names_to="data") %>%
  ggplot() +
  geom_point(aes(x=x, y=value), alpha=0.2) +
  facet_wrap(~data)

df %>%
  pivot_longer(-starts_with("x"), names_to="data") %>%
ggplot() +
  geom_histogram(aes(x=value, fill=data), alpha=0.8) +
  facet_wrap(~data)
```

# Reference Model

Fit `y_true` linear model as reference.

```{r results="hide"}
# mdl_data <- list(N = N,
#                  K = 2,
#                  y = df$y_true,
#                  x = matrix(c(df$x, df$x2), 
#                             nrow=N))

# Fit model
mdl1 <- stan(file=here("rounding_example/y_true.stan"), 
             data=list(N = N,
                 y = df$y_true,
                 x = df$x), 
             model_name="mdl1",
             chains=4)
```

```{r}
print(mdl1)
```

```{r}
mcmc_rank_overlay(mdl1)
```

```{r}
# Save posterior draws for comparison later
mdl1_draws <- mdl1 %>% 
  #gather_draws(alpha, beta[i], sigma) %>%
  gather_draws(alpha, beta, sigma) %>%
  mutate(model_id = "true")
```


# Rounded Data

## Incorrect Model

First, fit `y_round` data to the reference model to see effect of _not_ accounting for the rounding.

```{r results="hide"}
# Fit model
mdl2 <- stan(file=here("rounding_example/y_true.stan"), 
             data=list(N = N,
                 y = df$y_round,
                 x = df$x), 
             model_name="mdl2",
             chains=4)
```
```{r}
print(mdl2)
```

```{r}
mcmc_rank_overlay(mdl2)
```

```{r}
# Save posterior draws for comparison later
mdl2_draws <- mdl2 %>% 
  gather_draws(alpha, beta, sigma) %>%
  mutate(model_id = "round_wrong")
```
 
## Correct Model

Next, fit `y_round` data to the correct model twhich accounts for rounding.

```{r results="hide"}
# Fit model
mdl3 <- stan(file=here("rounding_example/y_round.stan"), 
             data=list(N = N,
                 y = df$y_round,
                 x = df$x), 
             model_name="mdl3",
             chains=4)
```

```{r}
print(mdl3)
```

```{r}
mcmc_rank_overlay(mdl3)
```

```{r}
# Save posterior draws for comparison later
mdl3_draws <- mdl3 %>% 
  gather_draws(alpha, beta, sigma) %>%
  mutate(model_id = "round_correct")
```

# Truncated Data

## Incorrect Model

Fit `y_trunc` data to the reference model to see effect of _not_ accounting for the truncating.

```{r results="hide"}
# Fit model
mdl4 <- stan(file=here("rounding_example/y_true.stan"), 
             data=list(N = N,
                 y = df$y_trunc,
                 x = df$x), 
             model_name="mdl4",
             chains=4)
```

```{r}
print(mdl4)
```

```{r}
mcmc_rank_overlay(mdl4)
```


```{r}
# Save posterior draws for comparison later
mdl4_draws <- mdl4 %>% 
  gather_draws(alpha, beta, sigma) %>%
  mutate(model_id = "trunc_wrong")
```

## Correct Model

Next, fit `y_trunc` data to the correct model which accounts for truncation.

```{r results="hide"}
# Fit model
mdl5 <- stan(file=here("rounding_example/y_trunc.stan"), 
             data=list(N = N,
                 y = df$y_trunc,
                 x = df$x), 
             model_name="mdl5",
             chains=4)
```

```{r}
print(mdl5)
```

```{r}
mcmc_rank_overlay(mdl5)
```


```{r}
# Save posterior draws for comparison later
mdl5_draws <- mdl5 %>% 
  gather_draws(alpha, beta, sigma) %>%
  mutate(model_id = "trunc_correct")
```


# Compare Models

```{r}
# Compare model fits
rbind(mdl1_draws, mdl2_draws,
      mdl3_draws, mdl4_draws,
      mdl5_draws) %>%
  group_by(.variable, model_id) %>%
  summarize(mean = mean(.value),
            lower = quantile(.value, probs=0.025),
            upper = quantile(.value, probs = 0.975),
            .groups="drop") %>%
  mutate(model_id = fct_relevel(model_id, "true")) %>%
  # arrange(.variable, model_id) %>%
  # knitr::kable()
  ggplot(aes(x=mean, xmin=lower, xmax=upper,
             y=model_id, color=model_id),) +
  geom_errorbarh(aes(height=0.1)) +
  geom_point() +
  facet_wrap(~.variable, scales="free_x") +
  labs(x="Posterior 95% Interval") +
  geom_vline(data = data.frame(xint=c(alpha_true, beta_true, sigma_true), 
                               .variable=c("alpha", "beta", "sigma")),
             aes(xintercept = xint), color="red",
             linetype="dashed") +
  theme(axis.text.x = element_text(angle = 45)) +
  theme(legend.position = "none")
```