---
title: Rounding & Truncating
date: \today
output:
  html_document:
    df_print: paged
  pdf_document: default
---

# Description

This file looks at the effects of rounding and truncation on parameter estimates for a simple linear model

$$y = \alpha + x\beta$$

There are three simulated data sets:

1.  y_true = raw data

2.  y_round = round(y_true)

3.  y_trunc = trunc(y_true)

Similarly there are three different models: *reference*, *rounded*, *truncated*.

A total of seven fits are performed:

1.  y_true with *reference* (This is the "reference" fit against which the others are compared).

2-4. y_round with *reference*, *rounded*, and *truncated*

5-7. y_trunc with *reference*, *rounded*, and *truncated*

# Setup

```{r libraries, results='hide', message=FALSE, warning=FALSE}

library(tidyverse)
library(RColorBrewer)
library(brms)
library(here)
library(bayesplot)
library(tidybayes)
```

```{r options}
options("scipen" = 1, "digits" = 4)

#knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(message = FALSE)
#knitr::opts_chunk$set(warning = FALSE)
#knitr::opts_chunk$set(out.width = "50%")
#knitr::opts_chunk$set(fig.align = "center")

# Set Default ggplot palette
options(ggplot2.discrete.color=brewer.pal(8, "Dark2"))
options(ggplot2.discrete.fill=brewer.pal(8, "Dark2"))

ncores <- min(4, parallel::detectCores()-1)

set.seed(1234)
```

# Simulate Data

```{r message=FALSE}
N <- 1000
alpha_true = -1
beta_true = 1
sigma_true = 1

df <- data.frame(x = rnorm(N, 0, 5)) %>%
  mutate(y_true = alpha_true + beta_true*x + rnorm(N, 0, sigma_true),
    y_round = round(y_true),
    y_trunc = trunc(y_true)
  )

df %>%
  pivot_longer(-starts_with("x"), names_to="data") %>%
  ggplot() +
  geom_point(aes(x=x, y=value), alpha=0.2) +
  facet_wrap(~data)

df %>%
  pivot_longer(-starts_with("x"), names_to="data") %>%
ggplot() +
  geom_histogram(aes(x=value, fill=data), alpha=0.8) +
  facet_wrap(~data)
```

# Reference Model

Fit `y_true` linear model as reference.

```{r results="hide"}

# Fit model
mdl1 <- brm(y_true ~ 1 + x,
            prior = c(set_prior("normal(0, 5)", class="Intercept"),
                      set_prior("exponential(1)", class="sigma")),
            data=df,
            family=gaussian(),
            chains=4,
            cores=ncores,
            file_refit = "on_change",
            file = here("rounding_example/mdl1.rds"))
```

```{r}
summary(mdl1)
```

```{r}
mcmc_rank_overlay(mdl1)
```

```{r}

# Save posterior draws for comparison later
mdl1_draws <- mdl1 %>%
  #gather_draws(alpha, beta[i], sigma) %>%
  gather_draws(b_Intercept, b_x, sigma) %>%
  mutate(model_id = "reference")
```


# Define custom brms functions

## Define Rounded Likelihood

```{r}
rounded_normal <- custom_family(
  "rounded_normal", 
  dpars = c("mu", "sigma"),
  lb = c(NA, 0),
  links = c("identity"), 
  type = "real"
)

stan_funs_rounded_normal <- "
  real rounded_normal_lpdf(real y, real mu, real sigma) {
    // Likelihood
    return log(Phi((y + 0.5 - mu)/sigma) - Phi((y - 0.5 - mu)/sigma));
  }
"
```

## Define Truncated Likelihood

```{r}
truncated_normal <- custom_family(
  "truncated_normal", 
  dpars = c("mu", "sigma"),
  lb = c(NA, 0),
  links = c("identity"), 
  type = "real",
)

stan_funs_truncated_normal <- "
  real truncated_normal_lpdf(real y, real mu, real sigma) {
    // Likelihood
    if(y < 0.0) {
       return log(Phi((y - mu) / sigma)
          - Phi((y - 1.0 - mu) / sigma));     
    }
    else {
      return log(Phi((y + 1.0 - mu) / sigma)
            - Phi((y - mu) / sigma));
    }
  }
"
```
# Rounded Data

## Incorrect Model (Reference)

First, fit `y_round` data to the reference model to see effect of *not* accounting for the rounding.

```{r results="hide"}
# Fit model
mdl2 <- brm(y_round ~ 1 + x,
            prior = c(set_prior("normal(0, 5)", class="Intercept"),
                      set_prior("exponential(1)", class="sigma")),
            data=df,
            family=gaussian(),
            chains=4,
            cores=ncores,
            file_refit = "on_change",
            file = here("rounding_example/mdl2.rds"))
```

```{r}
print(mdl2)
```

```{r}
mcmc_rank_overlay(mdl2)
```

```{r}
# Save posterior draws for comparison later
mdl2_draws <- mdl2 %>% 
  gather_draws(b_Intercept, b_x, sigma) %>%
  mutate(model_id = "round_reference")
```

## Correct Model



Next, fit `y_round` data to the correct model which accounts for rounding.

```{r results="hide"}
# Fit model
mdl3 <- brm(y_round ~ 1 + x,
            prior = c(set_prior("normal(0, 5)", class="Intercept"),
                      set_prior("exponential(1)", class="sigma")),
            data=df,
            family=rounded_normal,
            stanvars = stanvar(scode=stan_funs_rounded_normal,
                               block="functions"),
            #stan_funs = stan_funs_rounded_normal,
            chains=4,
            cores=ncores,
            file_refit = "on_change",
            file = here("rounding_example/mdl3.rds"))
```

```{r}
print(mdl3)
```

```{r}
mcmc_rank_overlay(mdl3)
```

```{r}
# Save posterior draws for comparison later
mdl3_draws <- mdl3 %>% 
  gather_draws(b_Intercept, b_x, sigma) %>%
  mutate(model_id = "round_round")
```

## Incorrect Model (Truncated)

First, fit `y_round` data to the truncated model to see effect of incorrect model.

```{r results="hide"}
# Fit model
mdl7 <- brm(y_round ~ 1 + x,
            prior = c(set_prior("normal(0, 5)", class="Intercept"),
                      set_prior("exponential(1)", class="sigma")),
            data=df,
            family=truncated_normal,
            stanvars = stanvar(scode=stan_funs_truncated_normal,
                               block="functions"),
            #stan_funs = stan_funs_rounded_normal,
            chains=4,
            cores=ncores,
            file_refit = "on_change",
            file = here("rounding_example/mdl7.rds"))
```

```{r}
print(mdl7)
```

```{r}
mcmc_rank_overlay(mdl7)
```

```{r}
# Save posterior draws for comparison later
mdl7_draws <- mdl7 %>% 
  gather_draws(b_Intercept, b_x, sigma) %>%
  mutate(model_id = "round_trunc")
```

# Truncated Data

## Incorrect Model (Reference)

Fit `y_trunc` data to the reference model to see effect of *not* accounting for the truncating.

```{r results="hide"}
# Fit model
mdl4 <- brm(y_trunc ~ 1 + x,
            prior = c(set_prior("normal(0, 5)", class="Intercept"),
                      set_prior("exponential(1)", class="sigma")),
            data=df,
            family=gaussian(),
            chains=4,
            cores=ncores,
            file_refit = "on_change",
            file = here("rounding_example/mdl4.rds"))
```

```{r}
print(mdl4)
```

```{r}
mcmc_rank_overlay(mdl4)
```

```{r}
# Save posterior draws for comparison later
mdl4_draws <- mdl4 %>% 
  gather_draws(b_Intercept, b_x, sigma) %>%
  mutate(model_id = "trunc_reference")
```

## Correct Model



Next, fit `y_trunc` data to the correct model which accounts for truncation.

```{r results="hide"}
# Fit model
mdl5 <- brm(y_trunc ~ 1 + x,
            prior = c(set_prior("normal(0, 5)", class="Intercept"),
                      set_prior("exponential(1)", class="sigma")),
            data=df,
            family=truncated_normal,
            stanvars = stanvar(scode=stan_funs_truncated_normal,
                               block="functions"),
            #stan_funs = stan_funs_rounded_normal,
            chains=4,
            cores=ncores,
            file_refit = "on_change",
            file = here("rounding_example/mdl5.rds"))
```

```{r}
print(mdl5)
```

```{r}
mcmc_rank_overlay(mdl5)
```

```{r}
# Save posterior draws for comparison later
mdl5_draws <- mdl5 %>% 
  gather_draws(b_Intercept, b_x, sigma) %>%
  mutate(model_id = "trunc_trunc")
```

## Incorrect Model (Rounding)

Next, fit `y_trunc` data to the model which assumes rounding.

```{r results="hide"}
# Fit model
mdl6 <- brm(y_trunc ~ 1 + x,
            prior = c(set_prior("normal(0, 5)", class="Intercept"),
                      set_prior("exponential(1)", class="sigma")),
            data=df,
            family=rounded_normal,
            stanvars = stanvar(scode=stan_funs_rounded_normal,
                               block="functions"),
            #stan_funs = stan_funs_rounded_normal,
            chains=4,
            cores=ncores,
            file_refit = "on_change",
            file = here("rounding_example/mdl6.rds"))
```

```{r}
print(mdl6)
```

```{r}
mcmc_rank_overlay(mdl6)
```

```{r}
# Save posterior draws for comparison later
mdl6_draws <- mdl6 %>% 
  gather_draws(b_Intercept, b_x, sigma) %>%
  mutate(model_id = "trunc_round")
```

# Compare Models

```{r}
# Compare model fits
rbind(mdl1_draws, mdl2_draws,
      mdl3_draws, mdl4_draws,
      mdl5_draws, mdl6_draws,
      mdl7_draws) %>%
  group_by(.variable, model_id) %>%
  summarize(mean = mean(.value),
            lower = quantile(.value, probs=0.025),
            upper = quantile(.value, probs = 0.975),
            .groups="drop") %>%
  mutate(model_id = fct_relevel(model_id, "true")) %>%
  # arrange(.variable, model_id) %>%
  # knitr::kable()
  ggplot(aes(x=mean, xmin=lower, xmax=upper,
             y=model_id, color=model_id),) +
  geom_errorbarh(aes(height=0.1)) +
  geom_point() +
  facet_wrap(~.variable, scales="free_x") +
  labs(x="Posterior 95% Interval") +
  geom_vline(data = data.frame(xint=c(alpha_true, beta_true, sigma_true), 
                               .variable=c("b_Intercept", "b_x", "sigma")),
             aes(xintercept = xint), color="red",
             linetype="dashed") +
  theme(axis.text.x = element_text(angle = 45)) +
  theme(legend.position = "none")
```


BLUF: We need to know whether data is rounded or truncated in order to use the correct model, otherwise the parameter estimates can be significantly biased.

The red dashed lines are the true parameter values. If the data is truncated, it is crucial that the _trucated_ model is used to get a correct estimate for all the parameters. If the data is rounded, the estimates are slightly more robust to model choice.  Obviously, the _rounded_ model yields the best estimates but the _reference_ model performs nearly as well.  However, using the _truncated_ model with rounded data gives incorrect parameter estimates.  
